WWII Events Reduce Population in One Shanghai District While Increasing It in Others

### 2024-12-03 paper/paper.qmd
added to appendix model testing

## Overview

@@

-   talk about where the data in my primary reference book is gathered
-   mentions the credibility of source and instability of recording instrument & methodology

@@

Place: Shanghai, China

When: 1936-1942

Who: Chinese population in Shanghai

Define: 1. give 1-2 sentence broad overview of China's state of unrest 2.THREE districts in Shanghai: Chinese District, International Settlement, French Concession - who controlled each district and the level of governance each authority have in comparison to Chinese government 3.Outline area (%) of each district (do not get into specifics, put that in Results section)

Important Event Timeline

@@

-   how population is recorded
-   Why choose 1936 - 1942 (ex. if an event occurred in Nov of 1937, would I take values of 1937 as a variable for prior to event occurrence or after?)

Survey Difficulties - Population Volatility caused by Warfare: Population is extreme volatile within this time period, population shifts constantly as 战况 proceed - Regional Recording Limitations: Survey response collect difficulty varies between districts. - Resistance due to Political Risk

@@

knitr::kable(combined_data, col.names = c("Year", "Pop CHN District", "Pop IS", "Pop FC"), align = c("c", "c", "c", "c"))
```

@@

-   outcome variable is *population*

@@

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

## Population Trend

```{r}
#| label: fig-pop
#| fig-cap: Population in Shanghai between 1936 - 1942
#| echo: false
#| warning: false
#| message: false

@@

## Interpretation of the Event Impacts

### Impact of Occupation in Chinese District

@@

# Additional data details

# Model details {#sec-model-details}

## Model Summary Table

library(rstanarm)

did_model <-
  readRDS(file = here::here("models/did_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "Population model" = did_model
  ),
  statistic = "mad",
  fmt = 2,
  metrics = c("R2", "RMSE")
)
```

@@

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

@@

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

>>>>

@@

The primary dataset is sourced from the appendix of the research publication 《旧上海人口变迁的研究》 (@sh), which includes selected excerpts from historical government census records. Due to the age of the original census documents, they are no longer available online. To address gaps in the primary dataset, complementary data was obtained from the research platform Virtual Shanghai (@virtualshanghai), providing crucial insights to fill missing areas.

@@

Between 1936 and 1942, China was in a state of political and social upheaval, marked by increasing Japanese aggression and internal instability. Shanghai, a critical urban center, was divided into three districts: the Chinese-administered district, the International Settlement (primarily governed by the U.K. and U.S.), and the French Concession (under French governance). The International Settlement and French Concession operated with significant autonomy compared to the Chinese government’s limited authority over its district. Each district occupied distinct proportions of Shanghai’s urban landscape.

**Reference Timeline**:

@@

The majority of the data used in this analysis were collected by police officers or hired local authorities within each district of Shanghai, as detailed in the source (@sh, page 2).

This paper focuses on population data from 1936 to 1942, a timeframe determined by both historical relevance and data availability. The analysis centers on the impact of three major historical events, beginning with the Japanese occupation of the Chinese-administered district on November 12, 1937. Given that Japanese forces first entered Shanghai on August 13, 1937, the year 1936 is chosen as the baseline or “no-treatment” year for comparison. The analysis concludes with the year 1942, as district-specific population data after this year is unavailable. The lack of records for the period between 1943 and 1949 coincides with the political transition to the People’s Republic of China, making 1942 the final year for which reliable district-based population data exists.

The survey process faced several challenges, primarily due to population volatility caused by wartime conditions. Population shifts were frequent and unpredictable as military conflicts unfolded. Additionally, survey efforts encountered varying degrees of difficulty across districts due to logistical and political constraints. Resistance from the local population, often rooted in political risks, further complicated data collection efforts. These challenges underline the complexities of accurately recording population dynamics during a period of significant unrest and upheaval.

@@

knitr::kable(combined_data, col.names = c("Year", "Pop CHN District", "Pop IS", "Pop FC"), align = c("c", "c", "c", "c"))
```
This paper primarily evaluates three historic events between 1936 - 1942:
1. The Japanese occupation of Shanghai's Chinese-administrated District -- 1937-11-12
2. France surrendered to Germany -- 1940-06-22
3. Japanese armed forces claims authority over International Settlement -- 1942-01

The first and third events should have very straight-forward consequences on population. The shifting over governing power to a foreign, hostile country would intuitive cause local population to panick and evacuate. In this paper, we initially assumed that Japanese occupation would decrease population in the corresponding area.

The effect of the second event is less direct. In 1940, European battlefield is still filled with WWII soldiers. When France surrendered to Germany in June, Japan -- an ally of Germany, accepted France's new status as mutural German allies. The Japanese forces chose to stay out of the French Concession District in Shanghai. (@video) The French Concession remained un-conquered until the Japanese withdrew from Shanghai in 1945.

@@

The outcome variable in this analysis is the *population* from 1936 to 1942.

District-level population data is missing for the years 1938, 1939, 1940 (partially), and 1941 in the government records. For 1938, 1939, and 1940, the population counts for all districts are unavailable due to undocumented reasons. To estimate these values, the population for the International Settlement and French Concession is deduced by adding the refugee population in their camps to their respective 1937 population counts. The population of the Chinese-administered district is then estimated by subtracting the total number of refugees in the two foreign districts from its 1937 population.

The missing population data for 1941 is estimated by calibrating the pre-established 1942 population dataset (@population1941) against the primary data. A multiplier is derived to adjust the 1941 data to align as closely as possible with the available records, providing an approximate but consistent estimation.

```{r}
#| echo: false
#| eval: true
#| label: tbl-pop
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false
#| message: false

knitr::kable(combined_data, 
             col.names = c("Year", "Chinese-administered Population", "International Settlement Population", "French Concession Population"), 
             align = c("c", "c", "c", "c"), 
             caption = "Population by District and Year (1936-1942)")

@@

### Model justification
The Difference-in-Difference model is chosen to evaluate the impact of historical events by comparing the observed outcomes, population changes, to an imaginary counterfactual scenario where these events did not occur.
The year variable captures the temporal trends, allowing the model to account for changes in population over time. The district variable identifies the three distinct regions in Shanghai (Chinese-administered district, International Settlement, and French Concession), each subject to unique governance and historical impacts. These inclusions enable the model to control for factors beyond the primary variables of interest (historical events).

A negative relationship is expected between year and population because the period under study (1936–1942) was marked by military conflict and foreign occupation, creating an unstable living environment. Specifically:
- Chinese-administered District: A population decline is expected due to the Japanese occupation in 1937.
- French Concession: A population decrease is anticipated following France’s surrender in 1940, which led to refugee rejection.
- International Settlement: A population decline is expected after its occupation by Japanese forces in 1942.

@@

## Population Trend
```{r}
#| label: fig-pop
#| fig-cap: Population in Shanghai between 1936 - 1942
#| echo: false
#| warning: false
#| message: false

@@

## Interpretation of the Event Impacts
The results indicate that the Japanese occupation of the Chinese-administered area led to an overall increase in Shanghai’s population across all districts. In contrast, France’s surrender during WWII had a negative impact on population numbers in all districts except the French Concession. Similarly, the Japanese occupation of the International Settlement caused population declines in all districts except the International Settlement itself.

### Impact of Occupation in Chinese District
- By intuition, we would think that the foreign-force occupation of a district would lead to its population to fled to other areas. However, this does not seem the case for the Chinese-administrated District within Shanghai.
- This counter-intuitive phenomenon might be caused by several factors. 
Firstly, the increase in the districts population might be cause by an influx of foreign state refugees. Shanghai was, and still is, one of the largest economic center in China and has an abundance of non-land-dependent job opportunity. 战争时期，随时随地可能发生的战事极大的interrupted agriculture, subsequently leading to a large volume of unemployed farmers. These unemployed population will likely flood to economic centers and industrial center in search of a new job. (@sh, page 14)

@@

# Model details {#sec-model-details}
## Model Summary Table

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Complete Model Summary for DiD model used in analysis"
#| warning: false

modelsummary::modelsummary(
  list(
    "Population model" = did_model
  ),
  statistic = "mad",
  fmt = 2,
  metrics = c("R2", "RMSE")
)
```
\newpage

@@

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. 
The posterior predictive check evaluates how well the model-generated predictions ($y_{rep}$) match the observed data ($y$). The overlay of the predictive distribution with the observed data indicates that the model captures the general trends and variability in the data. The similarity between the observed data and the model’s predictions suggests that the model provides a good fit for the observed population data.

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. 
The comparison between the posterior and prior distributions shows how the data influences the model parameters. The posterior distributions are narrower and more distinct compared to the priors, indicating that the data strongly informed the parameter estimates. Parameters like $cd\_occupied$ and $district\_fc$ show clear shifts in their posterior means, signifying their significant role in explaining population changes. Conversely, parameters with less distinct posteriors might indicate weaker evidence from the data.

@@

@fig-stanareyouokay-1 is a trace plot. The trace plot displays the MCMC sampling for each parameter across the chains. The chains appear to mix well, with no clear trends or patterns, indicating that the algorithm explores the parameter space efficiently. This suggests good convergence, as the chains reach their stationary distribution without being stuck in local modes.

@fig-stanareyouokay-2 is a Rhat plot. The Rhat plot evaluates the convergence of the chains by comparing the within-chain and between-chain variance. All Rhat values are close to 1.0, indicating that the chains have converged. This suggests that the parameter estimates are reliable, and the sampling process provides consistent results across chains.
